---
title: worminfo
author: "Michael J. Steinbaugh"
date: "`r Sys.Date()`"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "README-"
)
build <- get("build", envir = asNamespace("worminfo"))
numGenes <- nrow(get("geneSource", envir = asNamespace("worminfo")))
numClones <- nrow(get("rnaiSource", envir = asNamespace("worminfo")))
```

[Ahringer]: http://www.us.lifesciences.sourcebioscience.com/clone-products/non-mammalian/c-elegans/c-elegans-rnai-library/
[`devtools`]: https://cran.r-project.org/package=devtools
[Ensembl]: http://www.ensembl.org/Caenorhabditis_elegans
[ORFeome]: http://worfdb.dfci.harvard.edu
[PANTHER]: http://pantherdb.org
[R]: https://www.r-project.org
[WormBase]: http://www.wormbase.org

*C. elegans* genome annotations assembled from [WormBase][], [Ensembl][], and [PANTHER][]. RNAi clone mapping support for [ORFeome][], [Ahringer][] and cherrypick libraries.

[![Build Status](https://travis-ci.org/steinbaugh/worminfo.svg?branch=master)](https://travis-ci.org/steinbaugh/worminfo)


# Installation

This is an [R][] data package. [`devtools`][] is required to install the latest version directly from GitHub.

To install, run this code in [R][]:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("steinbaugh/worminfo", build_vignettes = TRUE)
```

Instructions on how to use the functions in this package are available as vignettes:

```{r, eval = FALSE}
browseVignettes("worminfo")
```


# Annotations

- `r numGenes` gene annotations
- `r numClones` RNAi clones

Compiled from these sources:

- `r build$wormbase`
- `r build$ensembl`
- `r build$panther`

Built with `r build$rstats`.
