---
title: worminfo
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: github_document
---

```{r data, include=FALSE}
devtools::load_all()
library(dplyr)
```

```{r, include=FALSE}
wormbase <- list.files(path = "data-raw/wormbase", pattern = "wormpep.table*") %>%
    stringr::str_extract(., "[0-9]{3}") %>%
    paste0("WormBase WS", .)
ensembl <- biomaRt::listMarts(host = "useast.ensembl.org") %>%
    dplyr::filter(biomart == "ENSEMBL_MART_ENSEMBL") %>%
    dplyr::select(version) %>%
    .[[1]]
panther <- list.files(path = "data-raw", pattern = "PTHR*") %>%
    stringr::str_extract(., "([0-9]{2}\\.[0-9]{1})") %>%
    paste0("PANTHER ", .)
rstats <- paste0(R.Version()$version.string,
                 " running on ",
                 R.Version()$platform)
```

[Ahringer]: http://www.us.lifesciences.sourcebioscience.com/clone-products/non-mammalian/c-elegans/c-elegans-rnai-library/
[Ensembl]: http://www.ensembl.org/Caenorhabditis_elegans
[ORFeome]: http://worfdb.dfci.harvard.edu
[PANTHER]: http://pantherdb.org
[R]: https://www.r-project.org
[WormBase]: http://www.wormbase.org

[![Build Status](https://travis-ci.org/steinbaugh/worminfo.svg?branch=master)](https://travis-ci.org/steinbaugh/worminfo)

<!--
[![codecov](https://codecov.io/gh/steinbaugh/worminfo/branch/master/graph/badge.svg)](https://codecov.io/gh/steinbaugh/worminfo)
-->

*C. elegans* genome annotations assembled from [WormBase][], [Ensembl][], and [PANTHER][]. RNAi clone mapping support for [ORFeome][], [Ahringer][] and cherrypick *E. coli* feeding libraries.

# Installation

This is an [R][] data package.

To install, run `devtools::install_github("steinbaugh/worminfo")`

# Annotations

- `r nrow(geneData)` gene annotations
- `r nrow(rnaiData)` RNAi clones

Compiled from `r wormbase`, `r ensembl` and `r panther` with `r rstats`.
