---
title: "worminfo"
output:
  md_document:
    variant: markdown_github
---

```{r, include=FALSE}
data(geneData)
data(rnaiData)
```

*C. elegans* genome and RNAi library annotations. Metadata dynamically assembled from [WormBase](http://www.wormbase.org), [Ensembl](http://www.ensembl.org/Caenorhabditis_elegans), and [PANTHER](http://pantherdb.org). Support for ORFeome and Ahringer feeding RNAi libraries.

[![Build Status](https://travis-ci.org/seqcloud/worminfo.svg?branch=master)](https://travis-ci.org/seqcloud/worminfo)

## Installation

This is an [R](https://www.r-project.org) data package.
Current version is `r packageVersion("worminfo")` (`r Sys.Date()`).

To install, run `devtools::install_github("seqcloud/worminfo")`.

## Annotations

- Genes: `r nrow(geneData)`.
- Ahringer clones: `r length(unique(na.omit(rnaiData$ahringer$orf)))`.
- ORFeome clones: `r length(unique(na.omit(rnaiData$orfeome$orf)))`.

```{r, echo=FALSE}
print(names(geneData))
```

Built with `r R.Version()$version.string` running on `r R.Version()$platform`.
